---
title: "Psych 1420, Assignment 2"
author: "YOUR NAME HERE!"
date: "DATE OF SUBMISSION HERE!"
output:
  html_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0.

## How to use R Markdown documents

This file is an R Markdown file. This is a special file that lets you write R code and run it, while also writing descriptive text (like this) about the code and analyses you have in between.

Every time you see a special colored block, like the one below, **that's where you can write and run R code.** This is called an **R code chunk.** You can type normal R code in the box, with _one command per line,_ and then run all the code in the box by pressing the green triangle (kind of looks like a play button) in the top right-hand corner of the chunk.

Try playing the code below to see what happens!

```{r}
my_variable <- 2 + 2
```

You should see in your Environment pane (top right-hand of RStudio window) now that you have a variable called `my_variable`, just as if you had typed that code into the console (bottom left-hand of RStudio window, where you did Assignment 1 in Swirl) and pressed Enter to run it there.

You can also use chunks to print out outputs in your document. These outputs will be included when you export the document later, so your TA will be able to see the results of all your R code.

For example, run the following code below using the green play button and see what happens:

```{r}
my_variable
```

Now, the value assigned to `my_variable` appears in the document. Handy! You'll be using this later to show the output of correlations and t-tests, as well as graphs you'll be making.

For every question in this assignment, every time you see an _empty_ code chunk, you will be expected to _type correct code into the chunk and run it._ We've already created code chunks for every question, so you can fill in the appropriate code in each section.

Every empty chunk you should fill in is marked with a **comment** that says "Code goes here". A comment is a special small note you can leave next to R code inside of a code chunk. You have to tell R that it's a comment, and not real code, by typing the pound/hashtag-sign # _before_ anything in your comment.

```{r}
# This will not run.
# If you press the green play button on this chunk, nothing should come out!
```

Any text after a # is not "seen" as code by R until the next line break. RStudio shows this kind of "comment text" in green (or another color, depending on your color scheme).

You can comment whole lines, if you want to type complete sentences in your notes. You can also add comments at the end of a line, if you want a comment to apply to a specific line of code.

```{r}
# The code BEFORE the # runs, and the text AFTER the # doesn't run.
2 + 2 # This should still output 4 because 2 + 2 is before the comment sign
```

We recommend only typing small "notes to self" in the comments inside code chunks. You can type anything longer than a couple sentences in the "normal text" part of your document, so that it will look like normal text when you export your document later.

## Packages

While R can do a lot of amazing stuff on its own, some special types of analysis or graphing can be more easily (or stylishly) done using one of a (huge) number of "packages" that people have created specially for R. Think of R as your phone, and the packages as apps that each allow the phone to do something extra.

To use a package, you have to install it (once), and then load it (every time you re-open RStudio).

In this assignment, we'll need the assistance of two packages in particular:

* `ggplot2` is the package we'll use to make clean, attractive graphs
* `knitr` is the package we'll use to export this homework assignment to a format you can upload to Canvas

First, you can use the code below to install these two packages if you haven't already.

```{r, eval = FALSE}
# Only run this code if you HAVEN'T already installed these packages!
install.packages(c("ggplot2",
                   "knitr"))
```

Next, use the following code to _load,_ or "turn on", these packages for the current RStudio window so you can actually access the functions inside. You need to do this every time you open a new RStudio window!

```{r}
# We only need to "turn on" ggplot2
# knitr actually works behind the scenes already!
library(ggplot2)
```

Now, we're ready to get started.

# 1.

Check your working directory using `getwd()`. If you opened RStudio via the R project file in the same directory that contains your datafile (`classdata_2018.csv`) and your R scripts, you should be in that folder now.

```{r}
# Code goes here

```

# 2.

Read in your data file using `read.csv("classdata_2018.csv")`, and assign it to the dataframe name `IntroSurvey`.

```{r}
# Code goes here

```

# 3.

Use the following exploration functions to check that the dataframe `IntroSurvey` read in correctly:

`head()`
 
```{r}
# Code goes here

```
 
`summary()`
 
```{r}
# Code goes here

```
 
`str()`
 
```{r}
# Code goes here

```

# 4.

Calculating variables from the Regret and Maximization scales:

Run the code below to create a new variable for overall Regret score in the `IntroSurvey` dataframe called `regret_total`. (This is the same `regret_total` that we calculated at the end of Assignment 1.)

```{r}
IntroSurvey$regret_total <- rowMeans(cbind(IntroSurvey$regret1,
                                           IntroSurvey$regret2,
                                           IntroSurvey$regret3,
                                           IntroSurvey$regret4,
                                           IntroSurvey$regret5))
```

You'll take a similar approach to calculating an overall Maximizing score from the six items we took from the short version of the Maximizing Scale. They are called `maxi1`, `maxi2`, `maxi3`, etc.

First, assign a new variable called `maxi_total` that is the average of all six `maxi` scores (if you need help, see the code above for calculating the mean regret score).

```{r}
# Code goes here

```

The Maximizing Scale has been shown to divide up into three subscales:

* Alternative Search (`maxi1` & `maxi3`)
* Decision Difficulty (`maxi2` & `maxi5`)
* High Standards (`maxi4` & `maxi6`)

Please assign three new variables in the `IntroSurvey` dataframe, each of which represents one subscale: `maxi_as`, `maxi_dd`, and `maxi_hs`. (Hint: you should be averaging the two scores for each.)

```{r}
# Code goes here

```

# 5.

Use `head()` to check that the following new variables are present in `IntroSurvey`:

* `regret_total`
* `maxi_total`
* `maxi_as`, `maxi_dd`, & `maxi_hs`

```{r}
# Code goes here

```

# 6.

Now we can start doing some real data analysis! For the purposes of getting practice in a variety of inferential statistics in R, we'll be running more exploratory tests than would normally be considered wise for a real study. We'll talk more about why this is a problem later in the semester, but hopefully in your stats class you talked about the dangers of running too many tests--i.e., alpha-inflation.

Let's start with some simple descriptive statistics on our new variables. Use `mean()` and `sd()` to find the mean and standard deviations for:

total regret

```{r}
# Code goes here

```

total maximizing

```{r}
# Code goes here

```

each of the maxi subscales. 

```{r}
# Code goes here

```

Use `hist()` to visualize the range of scores for each of the variables above.

```{r}
# Code goes here

```

```{r}
# Code goes here

```

```{r}
# Code goes here

```

Use `summary()` to get a summary of some of our _categorical_ variables (variables that are not numeric, but rather can take on one of a few specified categories):

`class`

```{r}
# Code goes here

```

`school`

```{r}
# Code goes here

```

`gender`

```{r}
# Code goes here

```

`handed`

```{r}
# Code goes here

```

Explore at least **5** other variables that you think might have a relationship with maximizing and/or regret. If the variable is _numeric_ or _interval_ (i.e., if its responses are numbers), compute the mean and sd, and produce a histogram. If the variable is _categorical_, produce a summary.

```{r}
# Code goes here

```

```{r}
# Code goes here

```

```{r}
# Code goes here

```

```{r}
# Code goes here

```

```{r}
# Code goes here

```

# 7.

Now that you have a sense for what the data show, let's look at how some of the variables relate to each other. First, some correlations.

## a.

Let's see if age correlates with birthyear (it should!). The function `cor.test()` allows us to put in two variables, see what their correlation is, and return a p-value for that correlation.

`cor.test()` allows you to run a Pearson, Kendall, or Spearman correlation, depending on what you tell it to do using the argument `method`. This time, since we just want to do a Pearson correlation, we'll use `method = "pearson"`.

First, we'll save the contents of our correlation test into the variable `cor_age_birthyear` so we can access its contents later. Then, we'll print `cor_age_birthyear` to the console, so we can visually inspect the results of the correlation test.

```{r}
cor_age_birthyear <- cor.test(IntroSurvey$age, IntroSurvey$birthyear, method = "pearson")

cor_age_birthyear
```

Above, you can see the r-value, p-value, and several other pieces of info about the correlation test. These statistics are actually saved as variables _inside of_ the correlation test object. To extract these statistics and report them, we can use the dollar sign `$` to index these statistics, similar to indexing columns in a dataframe.

To index the Pearson's r for this test, we can look for the sub-variable called `estimate`:

```{r}
cor_age_birthyear$estimate
```

And to index the p-value, we can call the sub-variable `p.value`:

```{r}
cor_age_birthyear$p.value
```

Now, we can use these variable names to refer to the Pearson's r and p-value of this correlation test. You don't have to look at the numbers and copy them over!

PS: In RStudio, if you want to know all the sub-variables inside of an object, if you just type the name of the object with a `$` after it, an auto-complete menu should pop up and you can use the up and down arrow keys to look through all the possible sub-variables.

## b.

Let's see if maximizing correlates with regret. Use `cor.test()` on those two variables, and report the Pearson's r and p-value.

```{r}
# Code goes here

```

```{r}
# Code goes here

```

```{r}
# Code goes here

```

Typically, maximizing correlates reasonably strongly with regret (r ~ 0.4). Does this appear to be the case in our dataset? 

## c.

Let's see if maximizing correlates with various measures of course-selection behavior. Calculate the correlation and report the Pearson's r between `maxi_total` and each of the following variables:

`courses_enrolled`

```{r}
# Code goes here

```

`courses_shopped`

```{r}
# Code goes here

```

`points_enrolled`

```{r}
# Code goes here

```

## d.

Is the regret score, as calculated by the Regret Scale, correlated with our students' self-reports of regret (`process_regret` and `outcome_regret`)?

```{r}
# Code goes here

```

```{r}
# Code goes here

```

What about `regret_general`?

```{r}
# Code goes here

```

## e.

Is the score on the Maximizing Scale related to the self-report of tendency to maximize (`maxi_general`)?

```{r}
# Code goes here

```
