---
title: "Untitled"
author: "Michelle.VanTieghem"
date: "2/3/2019"
output: html_document
---

```{r}
template <- read.csv("../assignment2revised/class_survey2_spring2019.csv")
var_names <- names(template)[1:43]
length(var_names)

names_MS <- var_names[7:19]

names_courses <- var_names[23:34]
names_MTS <- var_names[35:43]

```


```{r}
survey1 <- read.csv("survey1.csv")
survey2 <- read.csv("survey2.csv")

names(survey1)[2] <- "UNI"
names(survey2)[2] <- "UNI"

combined <- merge(survey1, survey2, by = "UNI")
names(combined)
ncol(combined)

# remove time stamp and uni 
library(tidyverse)
combined2 <- combined %>%
  select(-starts_with("Timestamp"), -starts_with ("UNI"),
         -starts_with("Please"))
ncol(combined2)
names(combined2)[1:13] <- names_MS

names(combined2)[19] <- "Courses_enrolled"
names(combined2)[20] <- "Courses_shopped"
names(combined2)[21] <- "Courses_considered"
names(combined2)[22:33] <- names_courses
names(combined2)[34:42] <- names_MTS

```

# make summary variables 

Happy: Course selection satisfaction 
  = mean Happy1 & Happy2 
```{r}
combined2$Happy <- rowMeans(cbind(combined2$Happy1, combined2$Happy2))

```
  
External: Use of external resources for decision-making process
  = mean External1 & External2 
```{r}
combined2$External <- rowMeans(cbind(combined2$External1, combined2$External2))

```

Time: Time spent deciding
  = mean Time1 & (8-Time2)
```{r}
time2_reverse <- 8-combined2$Time2
combined2$Time <- rowMeans(cbind(combined2$Time1,time2_reverse))

```

Info: Collecting course information 
  = mean Info1 & Info2 
```{r}
combined2$Info <- rowMeans(cbind(combined2$Info1, combined2$Info2))

```


ActvNum: Number of courses a student attended, but did not enroll in. 
  = Course2 – Course1 
```{r}
combined2$ActvNum <- rowMeans(cbind(combined2$Courses_shopped, combined2$Courses_enrolled))

```

PssvNum: Number of courses a student considered, but did not enroll in.
  = Course3 – Course1



```{r}
combined2PssvNum <- rowMeans(cbind(combined2$Courses_considered, combined2$Courses_enrolled))

```

# add ID 
```{r}

combined2$ID <- (1:nrow(combined2))
```

# save 
```{r}
write.csv(combined2, "class_survey2_spring2019.csv")

```