---
title: "Psych 1420, R Assignment on ANOVA"
author: "YOUR NAME HERE!"
date: "DATE OF SUBMISSION HERE!"
output:
  html_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## open class data 
```{r}
my_df <- read.csv("../classdata_2018.csv")
head(my_df)
```


## Recode categorical variables of interest
```{r}
library(tidyverse)
library(corrplot)

cor_df <- my_df %>%
  select(-time_planning, -dec_mode, -class, -major, -school, -concentration, -gender, -handed)


corrplot(cor(cor_df))


summary(my_df$school)

# let's simplify our lives by only having 2 school groups
my_df2 <- my_df %>%
  filter(school == "CC" | school == "GS")

summary(my_df2$gender)
# again, let's simplify our lives by only having m/f

my_df2 <- my_df2 %>%
  filter(gender == "F" | gender == "M")

```

## T-test with 2 groups (between subjects)
Interpret this output. Are there differences in Y based on X?
```{r}
t.test(outcome_regret ~ school, data = my_df2)

```

## One-way anova with 2 groups (between subjects)
School has 2 categories: CC and GS
We will test whether school (CC vs. GS) influences outcome regret. 
```{r}
one_way_anova <- aov(outcome_regret ~ school, data = my_df2)
summary(one_way_anova)
```
Interpret the output. Are there differences in Y based on X? 
Impt note: one-way anova with 2 groups is identical to a t-test! 





## One-way anova with 3 groups (between subjects)


## Two-way anova with 2 categorical variables (between subjects)
```{r}
two_way_anova <- aov(outcome_regret ~ school*gender, data = my_df2)
summary(two_way_anova)
```


### anova is really a linear model with categorical X variables!
check that the output is identical. 
```{r}
linear_model <- lm(outcome_regret ~ age*reader, data = my_df2)
anova(linear_model)
```

### this is useful if you want to run an ANCOVA:
include covariates (continuous X variables) to control for in your anova.
For example, I want to test the 2x2 interaction, controlling for age.
```{r}
linear_model2 <- lm(maxi_general ~ school*gender + age, data = my_df2)
anova(linear_model2)
```
